---
title: "Tilman's Resource Competition : 2 species 2 resources"
author: "Arthur Capet"
date: "June 15, 2017"
output: 
#  pdf_document:
#    toc: yes
#  slidy_presentation:
  github_document:
        toc: yes
bibliography: biblio.bib
urlcolor: blue
---

This script allows to visualize the dynamics of two species depending competing for two ressources[@TILMAN].
You might want to have a look on the [course notes](https://www.overleaf.com/read/krhfddzjxnqc) before going any further.

In this third sesssion, we are going to add a second species to the system. 
So we will have two resources and two species, and be able to test assumptions about cohabitation. 

```{r, results='hide', message=FALSE}
library("deSolve") # For solving differential equations
library("FME")     # Toolbox to play with model perturbation, sensitivity analysis, etc..
```
  
# Function and parameter definitions

1. Parameters for growth should now given for both species $N_1$ and $N_2$.

2. Remember that, in order to be callable for both species, the function `Growth` receives species-specifc parameter as an input argument.

```{r}
pars<-c(
  # Species N1
  mN1  = .1     ,  # mortality N1
  mu1 = .5      ,  # Max Growth 
  limN1R1 = 60  ,  # Half-Saturation R1 for N1
  limN1R2 = 10  ,  # Half-Saturation R2 for N1
  a11     = .6  ,  # Resource preference for R1, N1 [0-1]
  
  # Resources
  g1  = 60      ,  # Supply R1 (max R1 if no consumption)
  g2  = 40      ,  # Supply R2 (max R2 if no consumption)
  gT  = 3       ,  # Relaxation time towards max Conc

  # Initial conditions
  N1_0 = 10     ,  # Initial population N1
  R1_0 = 50     ,  # Initial stock R1
  R2_0 = 50     ,  # Initial stock R2

    # Simulation
  duration = 300,
  dt=.1
)
  # Ressource type
  #ftype="Essential"
  ftype="InteractiveEssential"
```

In order to be callable for both species, the function `Growth` has been built to receive species-specifc parameters as an input argument. We don't have to modify it.

```{r}
Growth<- function (R1,R2,Pp,hneed=F) {
  # Pp gives the species parameters 
  # * limR1
  # * limR2
  # * mu 
  # The 'with' function executes the code in {} with elements of the list (first argument)
  #    included as part of the local environment
  with (as.list(Pp), {
#    print(Pp)
    fR1 <- R1 / (R1 + limR1)
    fR2 <- R2 / (R2 + limR2)
 #   print('fr&,fré')
  #  print(fR1)
   # print(fR2)
  #  print(R2)
    if (ftype %in% c("Essential",
                      "InteractiveEssential",
                      "PerfectlySubstitutive",
                      "Complementary",
                      "Antagonistic") ){
    }else{
      print(' F type unknown, imposing Essential type' )
             ftype <- "Essential"                 
                      }
    #############
    # Essential #
    #############
    if (ftype=="Essential"){
    f <- mu * pmin(fR1,fR2)
    h1 <- a
    h2 <- (1-a)
    }
    #########################
    # Interactive Essential #
    #########################
        if (ftype=="InteractiveEssential"){
    f <- mu * fR1*fR2
    a<-.2
    h1 <- (a)
    h2 <- (1-a)
    }
    ##########################
    # Perfectly Substitutive #
    ##########################
        if (ftype=="PerfectlySubstitutive"){
    f <- mu * (R1+R2)/ ( R1+R2  + limR1+ limR2 )
    h1 <- R1/(R1+R2)
    h2 <- R2/(R1+R2)
        }
     
    #################
    # Complementary #
    #################
    if (ftype=="Complementary"){
     f <- mu * ((R1+R2+R1*R2/10)/ (R1+R2+R1*R2/10+limR1+limR2))
     h1 <- R1/(R1+R2)
     h2 <- R2/(R1+R2)
    }
    ################
    # Antagonistic #
    ################
    if (ftype=="Antagonistic"){
     f <- mu * ((R1+R2-R1*R2/80)/ (R1+R2-R1*R2/80+limR1+limR2))
     h1 <- R1/(R1+R2)
     h2 <- R2/(R1+R2)
     }
    #############
    # Switching #
    #############
    # f <- mu * pmax(R1,R2)/ (pmax(R1,R2)+   limR1+limR2 )
    # 
    # h1 <- R1/(R1+R2)
    # h2 <- R2/(R1+R2)
    #  if (R1>R2){
    #  h1 <- 1
    #  } else {
    #    h1 <- 0
    #  }
    #  h2<-1-h1
    
    if (hneed){
      return(c(f=f,h1=h1,h2=h2))
    } else {
      return(f)
    }
  })
}
```


The function providing the temporal derivatives has to be adapted. 

1. There is an additional state variables.
2. Growth shoud be called two times, once, for each species. 

```{r}
simpleg <- function (t, X, parms) {
  with (as.list(parms), {
    N1 <- X[1]
    R1 <- X[2]
    R2 <- X[3]
    
    # Return the growth rate and consumption vectors for N1
    pN1<-c( limR1 = limN1R1 ,
            limR2 = limN1R2 ,
            mu    = mu1     , 
            a     = a11     )
    
    # print(pN1)
    
    G1<-Growth(R1,R2,pN1,T)
    
    # print(G1)
    f1<-G1["f"]
    h11<-G1["h1"]
    h12<-G1["h2"]
    
    # Time derivatives
    dN1 <- N1 * (f1 - mN1)

    ## The following lines allows to include a stochastic term in the supply
    supply1 = (g1 - R1)/gT 
    supply2 = (g2 - R2)/gT 


    dR1 <-   supply1 - N1*f1*h11 
    dR2 <-   supply2 - N1*f1*h12 
    
    # print(dN1)
        # print(dR1)
            # print(dR2)
    
    # Return the time derivative
    return(list(c(dN1, dR1 , dR2)))
  })
}
```

# A first simulation

As before, start with a dynamic run and display the evolution of state variables. 

- How should the vector of initial conditions be exteneded ? 

```{r}
X0 <- with(as.list(pars),c(N1_0,R1_0,R2_0))  
times <- seq(0, pars["duration"], by = pars["dt"]) # output wanted at these time intervals

out <- ode(y = X0, times = times, func = simpleg, parms = pars,method = "euler")
colnames(out)<-c("time","N1","R1","R2") 
plot(out)
```

## PLOT 1 : Trajectories, Resource space and ZNGI

We'll use the steady simulation to illustrate competition, trajectories and equilibrium on the resource plane.
First let us compute the growth values over the resource plane for both $N_1$ and $N_2$.
We will only display ZNGI for both species (ie. no colored background).

```{r, space}
R1space <- seq(0,80, length=80)
R2space <- seq(0,80, length=80)

with (as.list(pars), {
  pN1 <<-c( limR1 = limN1R1 ,
          limR2 = limN1R2 ,
          mu    = mu1,
          a =a11  )
})
  
f1space <- outer(R1space,R2space,Growth,Pp=pN1)

# ZNGI for N1
contour(R1space ,R2space ,f1space,levels=as.vector(pars["mN1"]),col="red",lty = "dotted", labels="ZNGI 1",
        lwd = 3,
        vfont = c("sans serif", "plain"),
        labcex=1.5,
        xlab = "R1",
        ylab="R2")

# trajectories of the dynamic simulation
lines(out[,"R1"],out[,"R2"]  )
points(out[1,"R1"],out[1,"R2"],pch = 19)

# Equilibrium points and resource supply point
points(out[nrow(out),"R1"],out[nrow(out),"R2"],col='red',pch = 19)
points(pars["g1"],pars["g2"],col='blue',cex=1.5,bg='blue',pch=21)
```


## PLOT2 : perturbation on the supply point

We now the perturbate supply point value, as in TP2.

```{r, supply perturbation}


  fCRL<-function(parinit){
    parsl<-pars
    parsl[names(parinit)]<-parinit
    
    X0 <- with(as.list(parsl),c(N1_0,R1_0,R2_0))  
    
    out<- ode(y = X0, times = times, func = simpleg, parms = parsl,method="euler")
    colnames(out)<-c("time","N1","R1","R2")  

    points(parsl["g1"],parsl["g2"],col='blue',cex=1.5,bg='blue',pch=21)
    
    lines(c(parsl["g1"],out[length(times),"R1"]),c(parsl["g2"],out[length(times),"R2"]) , lty=2 )
    points(out[1,"R1"],out[1,"R2"],pch = 19)
    points(out[nrow(out),"R1"],out[nrow(out),"R2"],col='red',pch = 19)
    return(c("R1eq"=out[nrow(out),"R1"], "R2eq"=out[nrow(out),"R2"]))
  }
  
  parRange <- matrix(nr = 2, nc = 2, c(0, 0, 80,80 ,
                                       50,   50, 80,80 ),
                     dimnames = list(c("g1","g2"), c("min", "max")))
  parRange

  contour(R1space ,R2space ,f1space,levels=as.vector(pars["mN1"]),col="red",lty = "dotted", labels="ZNGI 1",
          lwd = 3,
          vfont = c("sans serif", "plain"),
          labcex=1.5,
          xlab = "R1",
          ylab="R2")


  CRL<-modCRL(fCRL,parRange=parRange,num = 20)

```
## PLOT3: perturbation on the supply point - Cohabitation ?  

Finally, we will plot the supply point with a certain color, according to the results of competition at equilibrium: Which species survives? Is cohabitation possible ? 

Since trajectories in the ressources plane does not provide information about species variables, we will use the color of the supply points to indicate different cases for cohabitations.

In the plot above, we added the supply point with the command : 
`points(parsl["g1"],parsl["g2"],col='blue',cex=1.5,bg='blue',pch=21)`

Here, instead of "blue", we will use a variable COL, which has to be given a value "blue" if both species survive, "red" if only N1 survives, "green" if only N2 survives, and "black" if none survives. 

As we've seen previously, it may take a long time for a species to actually reach a population of exactly zero. 
Therefore, as a criterion for "non-survival", we suggest to use the fact that the population is below a given thresold, for instance `TOL =0.01`, rather than being exactly equal to zero.


## Exercice 2 (Optionnal)

Can we say that heterogeneity in the system can enlarge the region of the ressource space that allows cohabitation ? 

To test this, insert a random term in the ressource supply (use function `runif`), with a new paramter `gnoise`.

1. Can you see a change in the region of the ressource plane where supply points allow coabitation (try different valur of `gnoise`)?

Note that to consider a proper comparison, the function `modCRL` allows to use a fixed set of 'perturbed parameters' rather than picking random values at each time. You may also consider a regular sampling of the ressource place for the supply point (see help of `?modCRL`).




# References


  