---
title: "Tilman's Resource Competition : 2 species 2 resources + Predation"
author: "Arthur Capet"
date: "June 15, 2017"
output: 
  #github_document:
    #toc : yes
  slidy_presentation:
    toc: yes
urlcolor: blue
---

This script allows to visualize the dynamics of two species depending competing for two ressources[@TILMAN].
You might want to have a look on the [course notes](https://www.overleaf.com/read/krhfddzjxnqc) before going any further.

```{r,results='hide', message=FALSE}
library("deSolve")
library("FME")
```
  
# Function and parameter definitions

Here the question was : What happens if a third species N3 predates on N1 and N2 ? 

* Define Parameters for N3
* Extend the size of the initial condition vector. 
* Modify the model function `simpleg`
* N1 and N2 are treated as limiting ressources for N3




```{r}
pars<-c(
  # Populations
  mN1  = .1     ,  # mortality N1
  mN2  = .15    ,  # mortality N2
  mN3  = .1    ,  # mortality N3
  # param for growth N1
  mu1 = .5      ,  # Max Growth 
  limN1R1 = 40  ,  # Half-Saturation R1 for N1
  limN1R2 = 40  ,  # Half-Saturation R2 for N1
  a11     = .6  ,  # Resource preference for R1, N1 [0-1]
  # param for growth N2
  mu2 = .5      ,  # Max Growth 
  limN2R1 = 10  ,  # Half-Saturation R1 for N2
  limN2R2 = 45  ,  # Half-Saturation R2 for N2
  a21     = .4  ,  # Resource preference for R1, N2 [0-1]
  # param for growth N3
  mu3 = 0.5      ,  # Max Growth for N3
  limN3N1 = 5  ,  # Half-Saturation N1 for N3
  limN3N2 = 5  ,  # Half-Saturation N2 for N3
  a31     = .6  ,  # Resource preference for R1, N2 [0-1]
  # Resources
  g1  = 80      ,  # Supply R1 (max R1 if no consumption)
  g2  = 80      ,  # Supply R2 (max R2 if no consumption)
  gT  = 10      ,  # Relaxation time towards max Conc
  # Initial conditions
  N1_0 = 10     ,  # Initial population N1
  N2_0 = 30     ,  # Initial population N2
  N3_0 = 5     ,  # Initial population N2
  R1_0 = 40     ,  # Initial stock R1
  R2_0 = 60     ,  # Initial stock R2
  # Simulation
  dt=.1
)

Growth<- function (R1,R2,Pp,ftype="Essential", hneed=F) {
  # Pp gives the species parameters 
  # * limR1
  # * limR2
  # * mu 
  # The 'with' function executes the code in {} with elements of the list (first argument)
  #    included as part of the local environment
  with (as.list(Pp), {
    fR1 <- R1 / (R1 + limR1)
    fR2 <- R2 / (R2 + limR2)
    
    if (ftype %in% c("Essential",
                      "InteractiveEssential",
                      "PerfectlySubstitutive",
                      "Complementary",
                      "Antagonistic") ){
    }else{
      print(' F type unknown, imposing Essential type' )
             ftype <- "Essential"                 
                      }
    #############
    # Essential #
    #############
    if (ftype=="Essential"){
    f <- mu * pmin(fR1,fR2)
    h1 <- a
    h2 <- (1-a)
    }
    #########################
    # Interactive Essential #
    #########################
        if (ftype=="InteractiveEssential"){
    f <- mu * fR1*fR2
    a<-.2
    h1 <- (a)
    h2 <- (1-a)
    }
    ##########################
    # Perfectly Substitutive #
    ##########################
        if (ftype=="PerfectlySubstitutive"){
    f <- mu * (R1+R2)/ ( R1+R2  + limR1+ limR2 )
    h1 <- R1/(R1+R2)
    h2 <- R2/(R1+R2)
        }
     
    #################
    # Complementary #
    #################
    if (ftype=="Complementary"){
     f <- mu * ((R1+R2+R1*R2/10)/ (R1+R2+R1*R2/10+limR1+limR2))
     h1 <- R1/(R1+R2)
     h2 <- R2/(R1+R2)
    }
    ################
    # Antagonistic #
    ################
    if (ftype=="Antagonistic"){
     f <- mu * ((R1+R2-R1*R2/80)/ (R1+R2-R1*R2/80+limR1+limR2))
     h1 <- R1/(R1+R2)
     h2 <- R2/(R1+R2)
     }
    #############
    # Switching #
    #############
    # f <- mu * pmax(R1,R2)/ (pmax(R1,R2)+   limR1+limR2 )
    # 
    # h1 <- R1/(R1+R2)
    # h2 <- R2/(R1+R2)
    #  if (R1>R2){
    #  h1 <- 1
    #  } else {
    #    h1 <- 0
    #  }
    #  h2<-1-h1
    
    if (hneed){
      return(c(f=f,h1=h1,h2=h2))
    } else {
      return(f)
    }
  })
}

simpleg <- function (t, X, parms) {
  with (as.list(parms), {
    N1 <- X[1]
    N2 <- X[2]
    N3 <- X[3]
    R1 <- X[4]
    R2 <- X[5]
    
    # Return the growth rate and consumption vectors for N1
    pN1<-c( limR1 = limN1R1 ,
            limR2 = limN1R2 ,
            mu    = mu1     , 
            a     = a11     )
    
    G1<-Growth(R1,R2,pN1,hneed=TRUE)
    
    f1<-G1["f"]
    h11<-G1["h1"]
    h12<-G1["h2"]
    
    # Return the growth rate and consumption vectors for N2
    pN2<-c( limR1 = limN2R1 ,
            limR2 = limN2R2 ,
            mu    = mu2     , 
            a     = a21     )
    
    G<-Growth(R1,R2,pN2,hneed=TRUE)
    f2<-G["f"]
    h21<-G["h1"]
    h22<-G["h2"]
    
    
    # Return the growth rate and consumption vectors for N2
    pN3<-c( limR1 = limN3N1 ,
            limR2 = limN3N2 ,
            mu    = mu3     , 
            a     = a31     )
    
    G<-Growth(N1,N2,pN3,hneed=TRUE)
    f3<-G["f"]
    h31<-G["h1"]
    h32<-G["h2"]
    
    # Time derivatives
    dN1 <- N1 * (f1 - mN1) - N3*f3*h31
    dN2 <- N2 * (f2 - mN2) - N3*f3*h32
    dN3 <- N3 * (f3 - mN3)
    dR1 <-  (g1-R1)/gT - N1*f1*h11 - N2*f2*h21
    dR2 <-  (g2-R2)/gT - N1*f1*h12 - N2*f2*h22
    
    # Return the time derivative
    return(list(c(dN1, dN2, dN3, dR1 , dR2)))
  })
}
```

# A first simulation

As before we start with a dynamic run
```{r}

X0 <- with(as.list(pars),c(N1_0,N2_0,N3_0,R1_0,R2_0))
times <- seq(0, 10000, by = pars["dt"]) # output wanted at these time intervals

out <- ode(y = X0, times = times, func = simpleg, parms = pars,method = "euler")
colnames(out)<-c("time","N1","N2","N3","R1","R2")
plot(out)
```

# Exploration of the resource space
We'll use the steady simulation to illustrate competition, trajectories and equilibrium on the resource plane.
First let us compute the growth values over hte resource plane for $N_1$ and $N_2$.
```{r}
#outsteady<-steady(y = X0, time=c(0,Inf),func = simpleg, parms = pars, method= "runsteady")
outsteady<-steady(y = X0, func = simpleg, parms = pars)#, method= "runsteady")
outs <- outsteady$y

names(outs)<-c("N1","N2","N3","R1","R2")

R1space <- seq(0,80, length=80)
R2space <- seq(0,80, length=80)

N1space <- seq(0,100, length=100)
N2space <- seq(0,100, length=100)

with (as.list(pars), {
  pN1 <<-c( limR1 = limN1R1 ,
          limR2 = limN1R2 ,
          mu    = mu1,
          a =a11  )
  pN2 <<-c( limR1 = limN2R1 ,
          limR2 = limN2R2 ,
          mu    = mu2,
          a =a11 )
  pN3 <<-c( limR1 = limN3N1 ,
          limR2 = limN3N2 ,
          mu    = mu3,
          a =a31 )
})
  
f1space <- outer(R1space,R2space,Growth,Pp=pN1)
f2space <- outer(R1space,R2space,Growth,Pp=pN2)
f3space <- outer(N1space,N2space,Growth,Pp=pN3)

contour(R1space ,R2space ,f1space,levels=as.vector(pars["mN1"]),col="red",lty = "dotted", labels="ZNGI 1",
        lwd = 3,
        vfont = c("sans serif", "plain"),
        labcex=1.5,
        xlab = "R1",
        ylab="R2")

# ZNGI for N2
contour(R1space ,R2space ,f2space,levels=as.vector(pars["mN2"]),col="red",lty = "dotted", labels="ZNGI 2",lwd = 3,vfont = c("sans serif", "plain"), labcex=1.5,add=T)

# trajectories of the dynamic simulation
lines(out[,"R1"],out[,"R2"]  )
points(out[1,"R1"],out[1,"R2"],pch = 19)

# Equilibrium points and resource supply point
points(out[nrow(out),"R1"],out[nrow(out),"R2"],col='red',pch = 19)
points(outs["R1"],outs["R2"],col='red',cex=2.5)
points(pars["g1"],pars["g2"],col='blue',cex=1.5,bg='blue',pch=21)



plot.new()
contour(N1space ,N2space ,f3space,levels=as.vector(pars["mN3"]),col="red",lty = "dotted", labels="ZNGI 3",
        lwd = 3,
        vfont = c("sans serif", "plain"),
        labcex=1.5,
        xlab = "N1",
        ylab="N2")

lines(out[,"N1"],out[,"N2"]  )
points(out[1,"N1"],out[1,"N2"],pch = 19)

```



  